Music Perception: Pilot Stimuli Analysis

### 2021-01-25 analysis/stimuli/stimuli_analysis.Rmd
add notes and fix ylim
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

# DVs
- IOIs (average tempo)
- Transition points
- CV (tempo variability)
- KOT (articulation)
- KV (dynamics)
- KV-diff (dynamics)

# Results
- 1. [Accuracy](https://github.com/atsukotominaga/perception-v1.0/blob/master/analysis/analysis.Rmd)
- 2. Exploratory analysis

In order to see which stimulus was recognised as teaching performances, we decided to run exploratory analysis. We plotted correlations between each DV and the percentage of judgement as "teaching".

@@

# IOIs (average tempo)
```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

@@

ggscatter(ioi, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

@@

ggscatter(ioi_art, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

@@

ggscatter(ioi_dyn, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

# Transition points
## All
```{r ioi-tra-all, echo = FALSE}
ioi_tra <- dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(MidFile)]

@@

ggscatter(ioi_tra, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

@@

ggscatter(ioi_tra_art, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

@@

ggscatter(ioi_tra_dyn, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

# CV (tempo variability)
## All
```{r cv, echo = FALSE}
cv <- dt_ioi[, .(N = .N, CV = sd(IOI)/mean(IOI)), by = .(MidFile)]

@@

ggscatter(cv, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```
## Articulation
```{r cv-art, echo = FALSE}
cv_art <- dt_ioi[Skill == "articulation", .(N = .N, CV = sd(IOI)/mean(IOI)), by = .(MidFile)]

@@

ggscatter(cv_art, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

@@

ggscatter(cv_dyn, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

# KOT (only articulation stimuli)
```{r kot, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time[row] - data_onset$time[row+1] # offset(n) - onset(n+1)
   }
}

@@

ggscatter(kot_legato, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```
## Staccato
```{r kot-staccato, echo = FALSE}
kot_staccato <- dt_kot[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]

@@

ggscatter(kot_staccato, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```
# KV (only dynamics stimuli)
## Forte
```{r vel-forte, echo = FALSE}
dt_vel <- data_onset
# assign Subcomponents
# for each note
dt_vel$Subcomponent <- NA
# Legato
for (phrase in 1:length(ls_legato_2)){
   for (note in 1:length(ls_legato_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "articulation" & dt_vel$Note == ls_legato_2[[phrase]][note]] <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
   for (note in 1:length(ls_staccato_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "articulation" & dt_vel$Note == ls_staccato_2[[phrase]][note]] <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte_2)){
   for (note in 1:length(ls_forte_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "dynamics" & dt_vel$Note == ls_forte_2[[phrase]][note]] <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano_2)){
   for (note in 1:length(ls_piano_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "dynamics" & dt_vel$Note == ls_piano_2[[phrase]][note]] <- "Piano"
   }
}

@@

ggscatter(vel_forte, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

@@

ggscatter(vel_piano, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

# KV-diff (dynamics)
## Forte to Piano
```{r vel-diff-forte, echo = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
# assign Interval
dt_vel_diff$Interval <- rep(1:66, nrow(dt_vel_diff)/66)

@@

ggscatter(vel_diff_ftop, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

@@

ggscatter(vel_diff_ptof, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE)
```

>>>>

## DVs
- IOIs (average tempo)
- Transition points
- CV (tempo variability)
- KOT (articulation)
- KV (dynamics)
- KV-diff (dynamics)

## Results
- 1. [Accuracy](https://github.com/atsukotominaga/perception-v1.0/blob/master/analysis/analysis.Rmd)

The accuracy results showed that it was not clear whether participants could detect teaching performances because the stimuli selection was not appropriate.

- 2. Exploratory analysis

In order to see which stimulus was recognised as teaching performances, we decided to run exploratory analysis. We plotted correlations between each DV and the percentage of judgement as "teaching" regardless of their original conditions (i.e., teaching vs. performing).

@@

# IOIs
- average tempo and teaching judgement

```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

@@

ggscatter(ioi, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

@@

ggscatter(ioi_art, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

@@

ggscatter(ioi_dyn, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

# Transition points
- tempo at transition points (e.g., legato to staccato / forte to piano) and teaching judgement

## All
```{r ioi-tra-all, echo = FALSE}
ioi_tra <- dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(MidFile)]

@@

ggscatter(ioi_tra, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

@@

ggscatter(ioi_tra_art, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

@@

ggscatter(ioi_tra_dyn, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

# CV (tempo variability)
- tempo variability and teaching judgement

## All
```{r cv, echo = FALSE}
cv <- dt_ioi[, .(N = .N, CV = sd(IOI)/mean(IOI)), by = .(MidFile)]

@@

ggscatter(cv, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Articulation
```{r cv-art, echo = FALSE}
cv_art <- dt_ioi[Skill == "articulation", .(N = .N, CV = sd(IOI)/mean(IOI)), by = .(MidFile)]

@@

ggscatter(cv_art, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

@@

ggscatter(cv_dyn, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

# KOT (only articulation stimuli)
- articulation and teaching judgement

```{r kot, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time[row] - data_onset$time[row+1] # offset(n) - onset(n+1)
   }
}

@@

ggscatter(kot_legato, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "KOT (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Staccato
```{r kot-staccato, echo = FALSE}
kot_staccato <- dt_kot[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]

@@

ggscatter(kot_staccato, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "KOT (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
# KV (only dynamics stimuli)
- dynamics and teaching judgement

## Forte
```{r vel-forte, echo = FALSE}
dt_vel <- data_onset
# assign Subcomponents
# for each note
dt_vel$Subcomponent <- NA
# Legato
for (phrase in 1:length(ls_legato_2)){
   for (note in 1:length(ls_legato_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "articulation" & dt_vel$Note == ls_legato_2[[phrase]][note]] <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
   for (note in 1:length(ls_staccato_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "articulation" & dt_vel$Note == ls_staccato_2[[phrase]][note]] <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte_2)){
   for (note in 1:length(ls_forte_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "dynamics" & dt_vel$Note == ls_forte_2[[phrase]][note]] <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano_2)){
   for (note in 1:length(ls_piano_2[[phrase]])){
     dt_vel$Subcomponent[dt_vel$Skill == "dynamics" & dt_vel$Note == ls_piano_2[[phrase]][note]] <- "Piano"
   }
}

@@

ggscatter(vel_forte, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "Velocity (0-127)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

@@

ggscatter(vel_piano, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "Velocity (0-127)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

# KV-diff (dynamics)
- dynamics changes at transition points (between forte and piano) and teaching judgement

## Forte to Piano
```{r vel-diff-forte, echo = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
# assign Interval
dt_vel_diff$Interval <- rep(1:66, nrow(dt_vel_diff)/66)

@@

ggscatter(vel_diff_ftop, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "Velocity Difference (0-127)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

@@

ggscatter(vel_diff_ptof, x = "Mean", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "Velocity Difference (0-127)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
